\input{template/lecture}
\input{template/macros}

% use tabs instead of spaces
% put references on slides

\subtitle{4. Feature Modeling}
\author{Elias Kuiter}

\begin{document}

\input{template/header}

\section{Feature Models}

% first: motivation and syntax of feature models/diagrams (how can a human model features sensibly?)
\input{content/modeling_notation}

\lessonslearned{
	\item Features, dependencies between features, configurators
	\item Feature models: abstract and concrete features, tree and cross-tree constraints
	\item Tree constraints: optional, mandatory, or group, alternative group
}{
	\item \fospl, Section~2.3 Feature Modeling
	\item Thorsten Berger et al. (2013): \href{https://doi.org/10.1145/2430502.2430513}{A Survey of Variability Modeling in Industrial Practice}
	\item Damir Nešić et al. (2019): \href{https://doi.org/10.1145/3338906.3338974}{Principles of Feature Modeling}
}{
	% \item In the feature model below, is the configuration $\{Base, API, Put, Trans., Win\}$ valid? % no, root+OS missing % block 3?
	% \featureDiagramDbmgmt

	% $\mathit{Transactions} \pimplies \mathit{Put} \por \mathit{Delete}$
	\item Sketch a feature model with at least 5 features (pen and paper preferred)
    \item Upload your solution to ...
    \item Discuss whether your solutions are syntactically correct
	% \item Sketch a feature model that has 9 valid configurations (pen and paper preferred)
 	% \item Discuss whether your solutions are syntactically correct and have the right number of configurations
}

%block 1: features explizit + ihre abhängigkeiten auch 
%notation für modell nutzen + eins zeichnen mit mind. 5 features, fehler korrigieren, lesen von modellen üben

\sectionend

% second: semantics of feature diagrams and other modeling notations (how can we encode features machine-readably?)
\section{Representations and Translations}

\input{content/modeling_representations}

\lessonslearned{
	\item To understand large configuration spaces, formal semantics and machine-readable representations matter.
	\item Propositional formulas satisfy many (though not all) needs for such a representation.
}{
	\item Don Batory (2005): \href{https://doi.org/10.1007/11554844_3}{Feature Models, Grammars, and Propositional Formulas}
	\item Alexander Knüppel et al. (2017): \href{https://doi.org/10.1145/3106237.3106252}{Is There a Mismatch Between Real-World Feature Models and Product-Line Research?}
}{
	\item Figure out how many valid configurations your feature diagram represents.
	When is this easy?
	\item Translate (a part of) your feature diagram into a formula, then into CNF, and finally into a DIMACS file.
	How might this be useful?
}
% translate your model into a propositional formula
%block 2: formel hinschreiben, valide + invalide cfg aufschreiben

\sectionend

% third: analyses of feature models (what can a computer tell us, which we cannot tell just by looking?)
\section{Automated Analyses}

\input{content/modeling_analyses}

\lessonslearned{
	\item Automated SAT-based analyses help in understanding large configuration spaces.
	\item There are still many open problems regarding encodings and applications (e.g., \#SAT).
}{
	\item David Benavides et al. (2010): \href{https://doi.org/10.1016/j.is.2010.01.001}{Automated Analysis of Feature Models 20 Years Later: A Literature Review}
	\item Thomas Thüm et al. (2009): \href{https://doi.org/10.1109/ICSE.2009.5070526}{Reasoning About Edits to Feature Models}
	\item Chico Sundermann et al. (2021): \href{https://doi.org/10.1145/3442391.3442404}{Applications of \#SAT Solvers on Feature Models}
}{
	% alternative: find inconsistencies in your feature diagram
	\item Recall that some $f \in F$ is dead iff $\pnot SAT(\phi \pand f)$.
	Suppose you want to know \emph{all} dead features of $\phi$.
	Naively, you may formulate $\abs{F}$ SAT queries.
	
	Can this be improved?
	
	(Hint: Some SAT solvers return an example model $M_\psi$ when $\psi$ is satisfiable.)
}

\input{template/footer}

\end{document}
