
\subsection{Software Configuration Management}

\begin{frame}{Excursus: Software Configuration Management}
	\begin{mycolumns}[columns=2,widths={35,63},animation=keep]
		\mydefinition{Software Configuration Management (SCM)}{	
			Policies, processes and tools for managing evolving software systems:
			\begin{itemize}
				\item Version control 
				\item System building
				\item Release management
				\item Change management
				\item Collaborative work
			\end{itemize}	
		}	
	\mynextcolumn
		\mydefinition{Basic Terms and Definitions}{	
			\begin{itemize}
				\item {\bf Software Item}: An (atomic) artifact that can be uniquely identified 
				\item {\bf Version}: A modified software item.
				\begin{itemize}
					\item {\bf Revision}: A new version that replaces an old one.
					\item {\bf Variant}: A version that co-exists with another one.
				\end{itemize}
				\item {\bf Configuration}: A set of software items that together form a functioning (partial) system.
				\item {\bf Baseline}: A stable configuration that represents a point of reference for further development.
				\item {\bf Release}: A baseline delivered to customers.				
			\end{itemize}
		}
	\end{mycolumns}	
\end{frame}

\begin{frame}{Excursus: Software Configuration Management}
	\vspace{2mm}
	\only<1|handout:0>{\pic[width=0.7\linewidth]{configuration-management-1}}%
	\only<2-|handout:1>{\pic[width=0.7\linewidth]{configuration-management-2}}%
\end{frame}

\begin{frame}{Example: A Conceptual Organization of our Graph Library}
	\vspace{2mm}
	\only<1|handout:0>{\pic[width=0.8\linewidth]{configuration-management-graphs-1}}%
	\only<2-|handout:1>{\pic[width=0.8\linewidth]{configuration-management-graphs-2}}%
\end{frame}

%\begin{frame}{Tool Support: Version Control Systems}
	%\only<1|handout:0>{\pic[width=\linewidth,page=1]{branching}}%
	%\only<2|handout:0>{\pic[width=\linewidth,page=2]{branching}}%
	%\only<3-|handout:1>{\pic[width=\linewidth,page=3]{branching}}%
%\end{frame}

\subsection{Version Control Systems}

\begin{frame}{Tool Support: Version Control Systems}
\begin{large}
	\hspace{70mm}
	\uncover<4->{$\text{cherrypick := patch}(\Delta(r8,r10),r11)$}\\
	\vspace{8mm}
	\only<1|handout:0>{\pic[width=0.8\linewidth]{versioncontrol-1}}%
	\only<2|handout:0>{\pic[width=0.8\linewidth]{versioncontrol-2}}%
	\only<3-|handout:1>{\pic[width=0.8\linewidth]{versioncontrol-3}}\\
	\vspace{8mm}
	\hspace{40mm}
	\uncover<4->{$\text{merge := 3-way-merge}(r4,\Delta(r4,r7),\Delta(r4,r9))$}
\end{large}
\end{frame}
% TODO no actual merges shown, only fast-forward

\subsection{Supporting Clone-and-Own Development}

\begin{frame}{Example: Graph Library under Version Control}
	\pic[width=0.7\linewidth]{versioncontrol-graphs-1.pdf}
\end{frame}

\begin{frame}{Example: Graph Library under Version Control}
	\pic[width=0.7\linewidth]{versioncontrol-graphs-2.pdf}
\end{frame}

\begin{frame}{Example: Graph Library under Version Control}
	\pic[width=0.7\linewidth]{versioncontrol-graphs-3.pdf}
\end{frame}

\subsection{Discussion}

\begin{frame}{Clone-and-Own with Version Control}
	\begin{mycolumns}[columns=2,widths={50,50},animation=keep]
		\mynote{Summary and Observations}{			
			Clone-and-Own with Provenance: % TODO provenance not introduced
			\begin{itemize}
				\item Supports keeping tracking of revisions and variants
				\item Creation of new variants is supported by merging of branches
				\item Propagation of changes between variants is supported by cherrypicking changes
			\end{itemize}	
			However:
			\begin{itemize}
				\item Versioning is typically limited to entire system variants (i.e., branches)
				\item No flexible combination of software items
			\end{itemize}
		}	
	\mynextcolumn
		\mynote{Advantages}{	
			\begin{itemize}
				\item Well-established and stable systems.
				\item Well-known known process.
				\item Good tool integration.	
			\end{itemize}
		}
		\mynote{Disadvantages}{	
			\begin{itemize}
				\item Development of variants, not features: flexible combination of features not directly possible.
				\item No structured reuse (copy \& edit).
				\item Merging and cherrypicking not fully automated.	
			\end{itemize}			
		}
	\end{mycolumns}	
\end{frame}